Reduced plasticity and relative variance in physiological rates of ectotherm populations under climate change

### 2024-06-18 ms.qmd
More responses and updating figures which are taking ages

@@

```{r, fig-1}
#| label: fig-1
#| fig-cap: "Taxonomic distribution of acute and acclimation $Q_{10}$ estimates across major habitats. **A)** Phylogenetic distribution of taxa contained within the data. The total number of acute and acclimation Q10 effect sizes are highlighted as well as whether the taxa are marine, freshwater or terrestrial. Silhouettes are representative taxa of major clades within the tree. **B)** Mean acute and acclimation $lnRR_{Q_{10}}$ across marine, freshwater, and terrestrial systems. **C)** Mean $lnCVR_{Q_{10}}$ across traits for marine, freshwater and terrestrial systems. Note there were no differences between acute and acclimation $Q_{10}$ types. k = total number of effect size estimates while the numbers in brackets indicate the number of species. Thick bars are 95% confidence intervals (CI) and thin bars 95% prediction intervals (PI)."
#| warning: false
#| echo: false

# Load and plot Figure 1. Note that this figure was derived using 3 separate code chunks described below. Figures were exported, imported into Adobe Illustrator for minor formatting changes and grid organisation. 
  fig1 <- image_read("./output/figures/fig1_final.png")
  fig1
##################
  # Figure 1a plot
##################
   # Map data to tree. Just one variable for now
      d <- ggtree(tree, layout="circular") %<+% tree_data

      # Plot the tree
      scales1 <- c("#053061", "#D1E5F0", "#67001F")
      scales2 <- c("#053061", "#E7E0DB", "#67001F")
      p1 <- d + 
            #geom_tiplab(size=2, offset = 10) + 
            geom_tippoint(aes(color=habitat), size = 2) +
            labs(color = "Habitat") + 
            scale_colour_manual(values = scales1)

      # Plot a heatmap of N for each trait around the phylogeny.
   phylo <-   gheatmap(p1, data.matrix(matrix_data2), offset = 0,  colnames = FALSE, 
               colnames_position="bottom", legend_title = "N", 
               colnames_angle=360, colnames_offset_y = 0, colnames_offset_x = -1.5, 
               hjust=0, font.size=3, width = 0.35) + 
                scale_fill_viridis(alpha = 0.6) + labs(fill="# of Effects")

##################
  # Figure 1b plot
##################
  viridis_cols <- c( "#21908CFF", "#440154FF","#228C22")

  k_lnrr <- manual_extract(model1.2, mod = "habitat", group = "record_num", exclude = NULL)

p1.2 <- orchaRd::orchard_plot(model1.2, mod = "habitat", group  = "record_num", at = list(type = c("acute", "acclim")), by = "type", weights = "prop", xlab = TeX("$lnRR_{Q_{10}}$"), angle = 45, legend.pos = "top.left", trunk.size = 1, branch.size = 2) + scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) +
  theme(legend.direction = "vertical") +

  annotate("text", x = 1.5, y = 0, 
         label =TeX(paste0("$\\beta = $", round(model1.2$b[3], 2), "[", round(model1.2$ci.lb[3], 2)," to ", round(model1.2$ci.ub[3], 2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 1.20, y = 1.4, 
         label =paste0(round(mean(change_f), 3), "%"), hjust = 0, size = 6) + labs(shape = "Type") +
  annotate("text", x = 1.20, y = -0.8, label = "P < 0.001", hjust = 0, size = 6) +

  annotate("text", x = 2.5, y = 0, 
         label =TeX(paste0("$\\beta = $", round(mean(m_acclim - m_acute), 2), "[", round(quantile(m_acclim - m_acute, 0.025), 2)," to ", round(quantile(m_acclim - m_acute, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 2.20, y = 1.4, 
         label =paste0(round(mean(change_m), 3), "%"), hjust = 0, size = 6) +
  annotate("text", x = 2.20, y = -0.8, label = "P = 0.01", hjust = 0, size = 6) +

  annotate("text", x = 3.5, y = 0, 
         label =TeX(paste0("$\\beta = $", round(mean(t_acclim - t_acute), 2), "[", round(quantile(t_acclim - t_acute, 0.025), 2)," to ", round(quantile(t_acclim - t_acute, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 3.20, y = 1.4, 
         label =paste0(round(mean(change_t), 3), "%"), hjust = 0, size = 6) +
  annotate("text", x = 3.20, y = -0.8, label = "P = 0.36", hjust = 0, size = 6) +

  scale_fill_manual(values = viridis_cols) + ylim(-0.5, 1.5) +  annotate("text", label= paste0("k = ", k_lnrr$k, " (", k_lnrr$sdy, ")"), x = c(0.8, 1.8, 2.8), y = 1.7, size = 5)

  p1.2 <- p1.2 + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 24),
                                       axis.text.y = element_text(angle = 45, size = 24),
                                       legend.title = element_text(size = 24),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 24))

  ggsave(p1.2, filename = "./output/figures/fig1.png", width = 9.1, height = 9.2, device = "png")

@@

  viridis_cols <- c( "#21908CFF", "#440154FF","#228C22")

@@

  p2 <- orchard_plot(model2.2, xlab = TeX("$lnCVR_{Q_{10}}$"), by = "type", at = list(type = c("acute", "acclim")), mod = "habitat", group  = "record_num", weights = "prop", angle = 45, trunk.size = 1, branch.size = 2, k.pos = "none") +
              scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) + ylim(-1.5, 1.5) + scale_fill_manual(values = viridis_cols) +

@@

  annotate("text", x = x3, y = -2, 
           label =TeX(paste0("$\\mu = $", round(mean(t_cvr_mean), 3), "[", round(quantile(t_cvr_mean, 0.025), 2)," to ", round(quantile(t_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = x3-0.15, y = 1, 
           label =paste0(round(mean(exp(t_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + labs(shape = "Type") +
    annotate("text", x = x3-0.15, y = -1.5, label = TeX(paste0("$p_{MCMC} = $", round(pmcmc(t_cvr_mean), digit = 3))), hjust = 0, size = 6) + ylim(-1.5, 1.75) + 
    annotate("text", label= paste0("k = ", k_cvr$k, " (", k_cvr$sdy, ")"), x = c(0.8, 1.8, 2.8), y = 1.3, size = 5) + ylim (-2, 2)

  p2 <- p2 + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 24),
                                       axis.text.y = element_text(angle = 45, size = 24),
                                       legend.title = element_text(size = 24),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 24))

@@

@@

### *Reduced variation in physiological rates in terrestrial and marine ectotherms*

@@

>>>>

```{r, fig-phylo}
#| label: fig-phylo
#| echo: false
#| fig-cap: "Phylogenetic distribution of acute and acclimation $Q_{10}$ estimates across major habitats.  The total number of acute and acclimation $Q_{10}$ effect sizes are highlighted as well as whether the taxa are marine, freshwater or terrestrial. Silhouettes are only representative taxa of major clades within the tree.

##################
  # Figure 2 plot
  # Phylogeny
##################
   # Map data to tree. Just one variable for now
      d <- ggtree(tree, layout="circular") %<+% tree_data

      # Plot the tree
      scales1 <- c("#053061", "#D1E5F0", "#67001F")
      p1 <- d + 
            #geom_tiplab(size=2, offset = 10) + 
            geom_tippoint(aes(color=habitat), size = 2) +
            labs(color = "Habitat") + 
            scale_colour_manual(values = scales1)

      # Plot a heatmap of N for each trait around the phylogeny.
   phylo <-   gheatmap(p1, data.matrix(matrix_data2), offset = 0,  colnames = FALSE, 
               colnames_position="bottom", legend_title = "N", 
               colnames_angle=360, colnames_offset_y = 0, colnames_offset_x = -1.5, 
               hjust=0, font.size=3, width = 0.35) + 
                scale_fill_viridis(alpha = 0.6) + labs(fill="# of Effects") + theme(legend.text =element_text(size = 24),
                                                                                    legend.title = element_text(size = 30, face = "bold"))

# Load and plot Figure 1. Note that this figure was derived using 3 separate code chunks described below. Figures were exported, imported into Adobe Illustrator for minor formatting changes and grid organisation. 
  fig1 <- image_read("./output/figures/fig1_final.png")
  plot(fig1)

@@

```{r, fig-1}
#| label: fig-1
#| fig-cap:  **A)** Mean acute and acclimation $lnRR_{Q_{10}}$ across marine, freshwater, and terrestrial systems. **C)** Mean acute and acclimation $lnCVR_{Q_{10}}$ across traits for marine, freshwater and terrestrial systems. Note there were no differences between acute and acclimation $Q_{10}$ types. For both plots, k = total number of effect size estimates while the numbers in brackets indicate the number of species. Sample sizes are the same for panel A and B. Thick bars are 95% confidence intervals (CI) and thin bars 95% prediction intervals (PI). $\Beta$ values are the contrasts between acute and acclimation means. $\mu$ values are the overall mean $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ averaged across acute and acclimdation types within each habitat type. In both cases, their 95% CI's are indicated within square brackets. $p_{MCMC}$ values are the posterior probability of the contrast being different from zero."
#| warning: false
#| echo: false

##################
  # Figure 2a plot
##################

  cols <- c("#053061","#D1E5F0",  "#67001F")
  k_lnrr <- manual_extract(model1.2, mod = "habitat", group = "record_num", exclude = NULL)

  p1.2 <- orchaRd::orchard_plot(model1.2, mod = "habitat", group  = "record_num", at = list(type = c("acute", "acclim")), by = "type", weights = "prop", xlab = TeX("$lnRR_{Q_{10}}$"), angle = 45, legend.pos = "top.left", trunk.size = 1.8, branch.size = 2.8, alpha = 0.4) + scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) +
  theme(legend.direction = "vertical") +

  annotate("text", x = 1.5, y = 0, 
         label =TeX(paste0("$\\beta = $", round(model1.2$b[4], 2), "[", round(model1.2$ci.lb[4], 2)," to ", round(model1.2$ci.ub[4], 2), "] ", " $p_{MCMC}$ < 0.001")), hjust = 0, size = 8) + 
  annotate("text", x = 1.20, y = -0.5, 
         label =paste0(round(mean(change_f), 3), "%"), hjust = 0, size = 8) + labs(shape = "Type") +

  annotate("text", x = 2.5, y = 0, 
         label =TeX(paste0("$\\beta = $", round(mean(m_acclim - m_acute), 2), "[", round(quantile(m_acclim - m_acute, 0.025), 2)," to ", round(quantile(m_acclim - m_acute, 0.975),2), "] ", " $p_{MCMC}$ = 0.01")), hjust = 0, size = 8) + 
  annotate("text", x = 2.20, y = -0.5, 
         label =paste0(round(mean(change_m), 3), "%"), hjust = 0, size = 8) +

  annotate("text", x = 3.5, y = 0, 
         label =TeX(paste0("$\\beta = $", round(mean(t_acclim - t_acute), 2), "[", round(quantile(t_acclim - t_acute, 0.025), 2)," to ", round(quantile(t_acclim - t_acute, 0.975),2), "] ", " $p_{MCMC}$ = 0.36")), hjust = 0, size = 8) + 
  annotate("text", x = 3.20, y = -0.5, 
         label =paste0(round(mean(change_t), 3), "%"), hjust = 0, size = 8) +

  scale_fill_manual(values = cols) + scale_color_manual(values = rep("black",3)) + ylim(-0.5, 1.5)

  p1.2 <- p1.2 + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 24),
                                       axis.text.y = element_text(angle = 45, size = 24),
                                       legend.title = element_text(size = 30),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 24), legend.position = "top")

@@

@@

  p2 <- orchard_plot(model2.2, xlab = TeX("$lnCVR_{Q_{10}}$"), by = "type", at = list(type = c("acute", "acclim")), mod = "habitat", group  = "record_num", weights = "prop", angle = 45, trunk.size = 1.8, branch.size = 2.8, alpha = 0.4, k.pos = "none") +
              scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) + ylim(-1.5, 1.5) +

  scale_fill_manual(values = cols) + scale_color_manual(values = rep("black",3)) +

@@

  annotate("text", x = x3, y = -2, 
           label =TeX(paste0("$\\mu = $", round(mean(t_cvr_mean), 3), "[", round(quantile(t_cvr_mean, 0.025), 2)," to ", round(quantile(t_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = x3-0.15, y = 1, 
           label =paste0(round(mean(exp(t_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + labs(shape = "Type") +
    annotate("text", x = x3-0.15, y = -1.5, label = TeX(paste0("$p_{MCMC} = $", round(pmcmc(t_cvr_mean), digit = 3))), hjust = 0, size = 6) + ylim(-1.5, 1.75) + 
    annotate("text", label= paste0("k = ", k_cvr$k, " (", k_cvr$sdy, ")"), x = c(0.8, 1.8, 2.8), y = 1.3, size = 5) + ylim (-1.5, 1.5)

  p2 <- p2 + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 24),
                                       axis.text.y = element_blank(),
                                       legend.title = element_text(size = 30),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 24), legend.position = "top")

@@

  (p1.2 + p2) + plot_annotation(tag_levels = "A", tag_suffix = ")")

@@

### *Relative variation in physiological rates is reduced in terrestrial and marine ectotherms*

@@

# Check the environmental variability across habitats. What is the average SD in temps

envi_var <- lnRRQ10_data_wild %>% group_by(habitat) %>% summarise(mean_SD = mean(sd), 
                                                                  mean_pred = mean(acf_all),
                                                                  stdy = length(unique(record_num)), 
                                                                        n = n())

###################################################
  # model_envi: Check how tempeartures vary by habitat
  # Expect terrestrial and freshwater systems tp be more variable 
  # Than marine. Predcition for autocorrelation is less clear
###################################################
# Lets have a look at modelling whether the environmental variability and predcitability vary by habitats
         sd_mod <- bf(log(sd) ~ habitat  + (1|record_num) + (1|spp)) 
    acf_all_mod <- bf(acf_all ~ habitat  + (1|record_num) + (1|spp))

model_envi <- brms::brm(sd_mod + acf_all_mod + set_rescor(TRUE), family = gaussian(),  control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 4, data = lnRRQ10_data_wild)
